---
title: "Ch 04 R Basics"
output:
  html_document:
    df_print: paged
---


## Exercises 

>1. What is the sum of the first 100 positive integers? The formula for the sum of integers $1$ through $n$ is $n(n+1)/2$. Define $n=100$ and then use R to compute the sum of $1$ through $100$ using the formula. What is the sum?

```{r}
n<-100
sum(n*(n+1)/2)
```


>2. Now use the same formula to compute the sum of the integers from 1 through 1,000.

```{r}
n<-1000
sum(n*(n+1)/2)
```

>3. Look at the result of typing the following code into R:
>
>```{r, eval=FALSE}
>n <- 1000
>x <- seq(1, n)
>sum(x)
>```
>
>    Based on the result, what do you think the functions `seq` and `sum` do?  You can use the `help` system:
>
>    A. `sum` creates a list of numbers and `seq` adds them up.
>
>    __B__. `seq` creates a list of numbers and `sum` adds them up.
>
>    C. `seq` computes the difference between two arguments and `sum` computes the sum of 1 through 1000.
>
>    D. `sum` always returns the same number.

```{r}
n<-1000
x<-seq(1,n)
sum(x)
```


>4. In math and programming, we say that we evaluate a function when we replace the argument with a given number. So if we type `sqrt(4)`, we evaluate the `sqrt` function. In R, you can evaluate a function inside another function. The evaluations happen from the inside out.  Use one line of code to compute the log, in base 10, of the square root of 100.

```{r}
log10(sqrt(100))
```


>5. Which of the following will always return the numeric value stored in `x`? You can try out examples and use the help system if you want.
>
>    A. $log(10^x)$
>
>    B. $log10(x^10)$
>
>    __C__. $log(exp(x))$
>
>    D. $exp(log(x, base = 2))$

```{r}
log(10^2)
log10(2^10)
log(exp(2))
exp(log(2, base = 2))
```

