---
title: "Ex 08 Data Visualization"
output:
  html_document:
    df_print: paged
    number_sections: true
    toc: true
    pandoc_args: [
      "--number-sections",
      "--number-offset=08"
    ]
---




## Exercises Distributions: ggplot2

Start by loading the __dplyr__ and __ggplot2__ library as well as the `murders` and `heights` data.

```{r}
library(dplyr)
library(ggplot2)
library(dslabs)
data(heights)
data(murders)
```

1\. With __ggplot2__ plots can be saved as objects. For example we can associate a dataset with a plot object like this

```{r, eval = FALSE}
p <- ggplot(data = murders)
```

Because `data` is the first argument we don't need to spell it out

```{r, eval = FALSE}
p <- ggplot(murders)
```

and we can also use the pipe:

```{r, eval = FALSE}
p <- murders %>% ggplot()
```

What is class of the object `p`?

 
2\. Remember that to print an object you can use the command `print` or simply type the object. 
Print the object `p` defined in exercise one and describe what you see.

a. Nothing happens.
b. A blank slate plot.
c. A scatterplot.
d. A histogram.


3\. Using the pipe `%>%`, create an object `p` but this time associated with the `heights` dataset instead of the `murders` dataset. 

   

4\. What is the class of the object `p` you have just created?

   
5\. Now we are going to add a layer and the corresponding aesthetic mappings. For the murders data we plotted total murders versus population sizes. Explore the `murders` data frame to remind yourself what are the names for these two variables and select the correct answer. __Hint__: Look at `?murders`.

a. `state` and `abb`.
b. `total_murders` and `population_size`.
c. `total` and `population`.
d. `murders` and `size`.

6\. To create the scatterplot we add a layer with `geom_point`. The aesthetic mappings require us to define the x-axis and y-axis variables, respectively. So the code looks like this:

```{r, eval=FALSE}
murders %>% ggplot(aes(x = , y = )) +
  geom_point()
```

except we have to define the two variables `x` and `y`. Fill this out with the correct variable names.


7\. Note that if we don't use argument names, we can obtain the same plot by making sure we enter the variable names in the right order like this:

```{r, eval=FALSE}
murders %>% ggplot(aes(population, total)) +
  geom_point()
```

Remake the plot but now with total in the x-axis and population in the y-axis.


8\. If instead of points we want to add text, we can use the `geom_text()` or `geom_label()` geometries. The following code 

```{r, eval=FALSE}
murders %>% ggplot(aes(population, total)) + geom_label()
```

will give us the error message: `Error: geom_label requires the following missing aesthetics: label`

Why is this?

a. We need to map a character to each point through the label argument in aes.
b. We need to let `geom_label` know what character to use in the plot.
c. The `geom_label` geometry does not require  x-axis and y-axis values.
d. `geom_label` is not a ggplot2 command.


9\. Rewrite the code above to use abbreviation as the label through `aes`


10\. Change the color of the labels to blue. How will we do this?

a. Adding a column called `blue` to `murders`.
b. Because each label needs a different color we map the colors through `aes`.
c. Use the `color` argument in `ggplot`.
d. Because we want all colors to be blue, we do not need to map colors, just use the color argument in `geom_label`.


11\. Rewrite the code above to make the labels blue.


12\. Now suppose we want to use color to represent the different regions. In this case which of the following is most appropriate:

a. Adding a column called `color` to `murders` with the color we want to use.
b. Because each label needs a different color we map the colors through the color argument of `aes` .
c. Use the `color` argument in `ggplot`.
d. Because we want all colors to be blue, we do not need to map colors, just use the color argument in `geom_label`.


13\. Rewrite the code above to make the labels' color be determined by the state's region.

14\. Now we are going to change the x-axis to a log scale to account for the fact the distribution of population is skewed. Let's start by defining an object `p` holding the plot we have made up to now

```{r, eval=FALSE}
p <- murders %>% 
  ggplot(aes(population, total, label = abb, color = region)) +
  geom_label() 
```

To change the y-axis to a log scale we learned about the `scale_x_log10()` function. Add this layer to the object `p` to change the scale and render the plot.

15\. Repeat the previous exercise but now change both axes to be in the log scale.

16\. Now edit the code above to add the title "Gun murder data" to the plot. Hint: use the `ggtitle` function.

## Exercises Distributions: eCDFs

### Qn
>1\. In the `murders` dataset, the region is a categorical variable and the following is its distribution:

```{r,message=FALSE}
library(ggplot2)
library(magrittr)
library(dplyr)
library(dslabs)
ds_theme_set()
data(murders)
murders %>% group_by(region) %>% summarize(n = n()) %>% 
    mutate(Proportion = n/sum(n),region = reorder(region, Proportion)) %>%
    ggplot(aes(x=region, y=Proportion, fill=region)) + 
    geom_bar(stat = "identity", show.legend = FALSE) + xlab("")
```

>To the closest 5%, what proportion of the states are in the North Central region?

```{r}
murders %>% group_by(region) %>%
  summarize(n = n()) %>%
  mutate(Proportion = n/sum(n), 
         region = reorder(region, Proportion)) %>%
  ggplot(aes(x=region, y=Proportion, fill=region)) + 
  geom_bar(stat = "identity", show.legend = FALSE) + 
  xlab("")
```


### Qn
>2\. Which of the following is true:
>
>a. The graph above is a histogram.
>b. The graph above shows only four numbers with a bar plot.
>c. Categories are not numbers, so it does not make sense to graph the >distribution.
>d. The colors, not the height of the bars, describe the distribution.


### Qn
>3\. The plot below shows the eCDF for male heights:
>
>
>```{r ecdf-exercise, echo=FALSE, eval = FALSE}
>heights %>% filter(sex == "Male") %>% ggplot(aes(height)) + 
>  stat_ecdf() +
>  ylab("F(a)") + xlab("a")
>```
>
>Based on the plot, what percentage of males are shorter than 75 inches?
>
>a. 100%
>b. 95%
>c. 80%
>d. 72 inches


### Qn
>4\. To the closest inch, what height `m` has the property that 1/2 of the male students are taller than `m` and 1/2 are shorter?
>    
>a. 61 inches
>b. 64 inches
>c. 69 inches
>d. 74 inches


### Qn
>5\. Here is an eCDF of the murder rates across states:
>
>```{r ecdf-exercise-2, echo=FALSE, eval = FALSE}
>murders %>% mutate(murder_rate = total/population * 10^5) %>%
>  ggplot(aes(murder_rate)) + 
>  stat_ecdf() +
>  ylab("F(a)") + xlab("a")
>```
>
>Knowing that there are 51 states (counting DC) and based on this plot, how many states have murder rates larger than 10 per 100,000 people?
>
>a. 1
>b. 5
>c. 10
>d. 50


### Qn
>6\. Based on the eCDF above, which of the following statements are true:
>
>a. About half the states have murder rates above 7 per 100,000 and the other half below.
>b. Most states have murder rates below 2 per 100,000.
>c. All the states have murder rates above 2 per 100,000.
>d. With the exception of 4 states, the murder rates are below 5 per 100,000.


### Qn
>7\. Below is a histogram of male heights in our `heights` dataset: 
>
>```{r height-histogram-exercise, echo=FALSE, eval = FALSE}
>heights %>% 
>  filter(sex == "Male") %>% 
>  ggplot(aes(height)) + 
>  geom_histogram(binwidth = 1, color = "black")
>```
>
>Based on this plot, how many males are between 63.5 and 65.5?
>
>a. 10
>b. 24
>c. 34
>d. 100


### Qn
>8\. About what **percentage** are shorter than 60 inches?
>  
>a. 1%
>b. 10%
>c. 25%
>d. 50%


### Qn
>9\. Based on the density plot below, about what proportion of US states have populations larger than 10 million?
>
>```{r density-exercise, echo=FALSE, eval = FALSE}
>murders %>% ggplot(aes(x=population/10^6)) + 
>  geom_density(fill = "grey") + 
>  scale_x_log10() +
>  xlab("Population in millions")
>```
>
>a. 0.02
>b. 0.15
>c. 0.50
>d. 0.55


### Qn
>10\. Below are three density plots. Is it possible that they are from the same dataset? 
>
>```{r density-exercise-2, echo=FALSE, warning=FALSE, message=FALSE, eval = FALSE}
>library(gridExtra)
>p1 <- murders %>% ggplot(aes(x=population/10^6)) + 
>  geom_density(fill = "grey", bw = 5) + xlab("Population in millions") + ggtitle("1")
>p2 <- murders %>% ggplot(aes(x=population/10^6)) + 
>  geom_density(fill = "grey", bw = .05) + scale_x_log10() + xlab("Population in millions") + ggtitle("2")
>p3 <- murders %>% ggplot(aes(x=population/10^6)) + 
>  geom_density(fill = "grey", bw = 1) + scale_x_log10() + xlab("Population in millions") + ggtitle("3")
>grid.arrange(p1,p2,p3,ncol=2)
>```
>
>Which of the following statements is true:
>
>a. It is impossible that they are from the same dataset.
>b. They are from the same dataset, but the plots are different due to code errors.
>c. They are the same dataset, but the first and second plot undersmooth and the third oversmooths.
>d. They are the same dataset, but the first is not in the log scale, the second undersmooths, and the third oversmooths.

## Exercises Distributions: Plots

### Qn
>1\. Define variables containing the heights of males and females like this:
>
>```{r, eval=FALSE}
>library(dslabs)
>data(heights)
>male <- heights$height[heights$sex == "Male"]
>female <- heights$height[heights$sex == "Female"]
>```
>
>How many measurements do we have for each?



### Qn
>2\. Suppose we can't make a plot and want to compare the distributions side by side. We can't just list all the numbers. Instead, we will look at the percentiles. Create a five row table showing `female_percentiles` and `male_percentiles` with the 10th, 30th, 50th, 70th, & 90th percentiles for each sex. Then create a data frame with these two as columns.
>

### Qn
>3\. Study the following boxplots showing population sizes by country:
>
>```{r boxplot-exercise, echo=FALSE, message=FALSE, eval=FALSE}
>library(tidyverse)
>library(dslabs)
>ds_theme_set()
>data(gapminder)
>tab <- gapminder %>% filter(year == 2010) %>% group_by(continent) %>% select(continent, population)  
>tab %>% ggplot(aes(x=continent, y=population/10^6)) + 
>  geom_boxplot() + 
>  scale_y_continuous(trans = "log10", breaks = c(1,10,100,1000)) + ylab("Population in millions")
>```
>
>Which continent has the country with the biggest population size?


### Qn
>4\. What continent has the largest median population size?
>    

### Qn 
>5\. What is median population size for Africa to the nearest million? 
>    

### Qn
>6\. What proportion of countries in Europe have populations below 14 million?
>    
>a. 0.99
>b. 0.75
>c. 0.50
>d. 0.25


### Qn
>7\. If we use a log transformation, which continent shown above has the largest interquartile range?

### Qn
>8\. Load the height data set and create a vector `x` with just the male heights:
>
>```{r, eval=FALSE}
>library(dslabs)
>data(heights)
>x <- heights$height[heights$sex=="Male"]
>```
>
>What proportion of the data is between 69 and 72 inches (taller than 69, but shorter or equal to 72)? Hint: use a logical operator and `mean`.

### Qn
>9\. Suppose all you know about the data is the average and the standard deviation. Use the normal approximation to estimate the proportion you just calculated. Hint: start by computing the average and standard deviation. Then use the `pnorm` function to predict the proportions.

### Qn
>10\. Notice that the approximation calculated in question nine is very close to the exact calculation in the first question. Now perform the same task for more extreme values. Compare the exact calculation and the normal approximation for the interval (79,81]. How many times bigger is the actual proportion than the approximation?

### Qn
>11\. Approximate the distribution of adult men in the world as normally distributed with an average of 69 inches and a standard deviation of 3 inches. Using this approximation, estimate the proportion of adult men that are 7 feet tall or taller, referred to as _seven footers_. Hint: use the `pnorm` function.

### Qn
>12\. There are about 1 billion men between the ages of 18 and 40 in the world. Use your answer to the previous question to estimate how many of these men (18-40 year olds) are seven feet tall or taller in the world?

### Qn
>13\. There are about 10 National Basketball Association (NBA) players that are 7 feet tall or higher. Using the answer to the previous two questions, what proportion of the world's 18-to-40-year-old _seven footers_ are in the NBA?

### Qn
>14\. Repeat the calculations performed in the previous question for Lebron James' height: 6 feet 8 inches. There are about 150 players that are at least that tall.

### Qn
>15\. In answering the previous questions, we found that it is not at all rare for a seven footer to become an NBA player. What would be a fair critique of our calculations:
>
>a. Practice and talent are what make a great basketball player, not height.
>b. The normal approximation is not appropriate for heights.
>c. As seen in question 10, the normal approximation tends to underestimate the extreme values. It's possible that there are more seven footers than we predicted.
>d. As seen in question 10, the normal approximation tends to overestimate the extreme values. It's possible that there are fewer seven footers than we predicted.


## Exercises Distributions: Geometries

### Qn
>1\. Now we are going to use the `geom_histogram` function to make a histogram of the heights in the `height` data frame. When reading the documentation for this function we see that it requires just one mapping, the values to be used for the histogram. Make a histogram of all the plots. 
>
>What is the variable containing the heights?
>
>a. `sex`
>b. `heights`
>c. `height`
>d. `heights$height`

### Qn
>2\. Now create a ggplot object using the pipe to assign the heights data to a ggplot object. Assign `height` to the x values through the `aes` function.

### Qn
>3\. Now we are ready to add a layer to actually make the histogram. Use the object created in the previous exercise and the `geom_histogram` function to make the histogram.

### Qn
>4\. Note that when we run the code in the previous exercise we get the warning:
>`stat_bin()` using `bins = 30`. Pick better value with `binwidth`.`
>
>Use the `binwidth` argument to change the histogram made in the previous exercise to use bins of size 1 inch. 

### Qn
>5\. Instead of a histogram, we are going to make a smooth density plot. In this case we will not make an object, but instead render the plot with one line of code. Change the geometry in the code previously used to make a smooth density instead of a histogram.

### Qn
>6\. Now we are going to make a density plot for males and females separately. We can do this using the `group` argument. We assign groups via the aesthetic mapping as each point needs to a group before making the calculations needed to estimate a density.

### Qn
>7\. We can also assign groups through the `color` argument. This has the added benefit that it uses color to distinguish the groups. Change the code above to use color.

### Qn
>8\. We can also assign groups through the `fill` argument. This has the added benefit that it uses colors to distinguish the groups, like this:
>
>
>```{r, eval=FALSE}
>heights %>% 
>  ggplot(aes(height, fill = sex)) + 
>  geom_density() 
>```
>
>However, here the second density is drawn over the other. We can make the curves more visible by using alpha blending to add transparency. Set the alpha parameter to 0.2 in the `geom_density` function to make this change.

