---
title: "Ch 06 Importing Data"
output:
  html_document:
    df_print: paged
    number_sections: true
    toc: true
    pandoc_args: [
      "--number-sections",
      "--number-offset=5"
    ]
---

## Exercises

```{r}
library(tidyverse)
```

>1. Use the `read_csv` function to read each of the files that the following code saves in the `files` object: 
>
>```{r, eval=FALSE}
>path <- system.file("extdata", package = "dslabs")
>files <- list.files(path)
>files
>```

```{r}
path <- system.file("extdata", package="dslabs")
files <- list.files(path)
files
```

Read all the files in.
```{r}
mortality <- paste(path,'/',files[1]) %>% str_replace_all(" ","")
fertility <- paste(path,'/',files[2]) %>% str_replace_all(" ","")
life <- paste(path,'/',files[3]) %>% str_replace_all(" ","")
murders <- paste(path,'/',files[4]) %>% str_replace_all(" ","")
olive <- paste(path,'/',files[5]) %>% str_replace_all(" ","")
#read_csv(mortality)
read_csv(fertility)
read_csv(life)
read_csv(murders)
read_csv(olive)
```

>2. Note that the last one, the `olive` file, gives us a warning. This is because the first line of the file is missing the header for the first column. 
>
>    Read the help file for `read_csv` to figure out how to read in the file without reading this header. If you skip the header, you should not get this warning. Save the result to an object called `dat`.

```{r}
olive_dat <- read.csv(olive,header=FALSE)
olive_dat %>% head
```

>3. A problem with the previous approach is that we don't know what the columns represent. Type:
>
>```{r, eval=FALSE}
>names(dat)
>```
>
>    to see that the names are not informative.
>
>    Use the `readLines` function to read in just the first line (we later learn how to extract values from the output). 

```{r}
olive_dat <- read.csv(olive,header=FALSE)
olive_dat %>% head
names(olive_dat)
olive_header <- readLines(olive,1) %>% str_split(',') %>% .[[1]]
olive_header[1] = 'Num'
olive_header[length(olive_header)+1] = 'sambalolic'
olive_header
#olive_dat <- setNames(olive_dat,olive_header)
#olive_dat %>% filter(Num!='NA')
```


