---
title: "Ex 15 Statistical Inference"
output:
  html_document:
    df_print: paged
    number_sections: true
    toc: true
    pandoc_args: [
      "--number-sections",
      "--number-offset=15"
      ]
---

## Exercises: Parameters Estimates

### Qn

>1\. Suppose you poll a population in which a proportion $p$ of voters are Democrats and $1-p$ are Republicans. Your sample size is $N=25$. Consider the random variable $S$ which is the **total** number of Democrats in your sample. What is the expected value of this random variable? Hint: it's a function of $p$.

The definition of the average is defined as the sum of its elements devided by the number of its elements. 

$$
\bar{X} = \frac{1}{N} \times \sum_{i=1}^N x_i = \frac{1}{N} X_N \Leftrightarrow N \bar{X} = X_N
$$
The expectation value is defined in the book as $\mbox{E}(\bar{X}) = p$. inserting the value $X=N\bar{X}$ into the standard Error one yields the following relation. 

$$
\mbox{E}(X) = N\ \mbox{E}(\bar{X}) = Np
$$
Similarly, for $S=N\bar{S}$ and $N=25$ one can write the following expression, where expected value is $25p$ denoting the number of Democrats for the sample size of 25 people. 
$$
\mbox{E}(S) = N\ \mbox{E}(\bar{S}) = Np = 25p
$$

### Qn
>2\. What is the standard error of $S$ ? Hint: it's a function of $p$.

Using the definition of the standard error of $\bar{X}$ as $\mbox{SE}(\bar{X}) = \sqrt{p(1-p)/N}$ and the definition $S=N\ \bar{S}$ one can write the following equation.

$$
\mbox{SE}(S) = N\ \mbox{SE}(\bar{S}) = N \sqrt{p(1-p)/N} = \sqrt{Np(1-p)} = \sqrt{25p(1-p)}
$$

### Qn
>3\. Consider the random variable $S/N$. This is equivalent to the sample average, which we have been denoting as $\bar{X}$. What is the expected value of the $\bar{X}$? Hint: it's a function of $p$.

Defining the random variable as $\bar{X} = S/N$ and utilizing the definition $\mbox{E}(S) = Np$ one can write the following equation.

$$
\mbox{E}(\bar{X}) = \mbox{E}(S/N) = \mbox{E}(S)\ \frac{1}{N} =  Np\ \frac{1}{N}= p
$$

### Qn
>4\. What is the standard error of $\bar{X}$? Hint: it's a function of $p$.

The standard error as a function of $\bar{X}$ is also provided in the book in chapter 15.2.4 "Properties of our estimate: expected value and standard error".
$$
\mbox{SE}(\bar{X}) = \sqrt{p(1-p)/N}
$$

### Qn
>5\. Write a line of code that gives you the standard error `se` for the problem above for several values of $p$, specifically for `p <- seq(0, 1, length = 100)`. Make a plot of `se` versus `p`. 

From this plot one can see the function $\mbox{SE[X(p)]}$ evolving by varying the parameter $p$. The function we see is symmetric. For probabilities p=0 and p=1 the standard error is zero, therefore for events with absolute certainty of occurring =1 or not occurring =0 there is no room for deviations. For the probability of p=0.5 the standard error is at its maximum, meaning that this point marks the biggest uncertainty for the probability. The standard error and therefore the uncertainty is falling off as p moves towards one of the certain points. 

```{r,message=FALSE}
library(tidyverse)
N <- 25
p<-seq(0,1,length=100)
se<-sqrt(p*(1-p)/N)
qplot(p,se)
````


### Qn
>6\. Copy the code above and put it inside a for-loop to make the plot for $N=25$, $N=100$, and $N=1000$. 

In the following plots it can be noticed that the standard error "se" gets smaller for increasing population size of N (from red to blue in the bottom plot).

```{r,message=FALSE}
vN<-c(25,100,1000)
p<-seq(0,1,length=100)
for(N in vN){
se<-sqrt(p*(1-p)/N)
plot(p,se,ylim = c(0,0.5/sqrt(25)))
}
````
```{r,message=FALSE}
library(tidyverse)
N25<-replicate(100,25)
N1000<-replicate(100,1000)
N100<-replicate(100,100)
N<-c(N25,N100,N1000)
p<-seq(0,1,length=100)
se<-sqrt(p*(1-p)/N)
df <- data.frame(p,se,N) %>% mutate(N=as.factor(N))
df %>% ggplot(aes(p,se,color=N)) + geom_point()
````

### Qn
>7\. If we are interested in the difference in proportions, $p - (1-p)$, our estimate is $d = \bar{X} - (1-\bar{X})$. Use the rules we learned about sums of random variables and scaled random variables to derive the expected value of $d$.

Here, the lower case letter $d$ denotes the spread.
$$
\mbox{E}(d) = \mbox{E}(\bar{X}) - (\mbox{E}(1)-\mbox{E}(\bar{X})) = p-(1-p) = 2p-1
$$

### Qn
>8\. What is the standard error of $d$?

The result from $\mbox{SE}(\bar{X})$ follows directly from the previous definition in the answer of question 4. However, the standard error of a constant value, here $\mbox{SE}(1)=0$ always equals to be 0, because there cannot be any deviation from a sample size of one element (the mean of the constant itself). In terms of equations the standard error can be defined utilizing the expectation value as $\mbox{SE}(\bar{X})^2=\mbox{E}(\bar{X^2})-\mbox{E}(\bar{X})^2$. For a constant value if follows $\mbox{SE}(1)^2=\mbox{E}(1^2)-\mbox{E}(1)^2 = 1-1=0$. 

$$
\mbox{SE}(d) = \mbox{SE}(\bar{X}) - (\mbox{SE}(1)-\mbox{SE}(\bar{X})) = \sqrt{p(1-p)/N}-(0-\sqrt{p(1-p)/N}) = 2\sqrt{p(1-p)/N}
$$


### Qn
>9\. If the actual $p=.45$, it means the Republicans are winning by a relatively large margin since $d= -.1$, which is a 10% margin of victory. In this case, what is the standard error of $2\hat{X}-1$ if we take a sample of $N=25$? 

The expectation value of the spread $d$ is 10%, where the standard error is about 20%. Therefore, the estimated error is double in size than the actual measurement. 

```{r,message=FALSE}
N<-25
p<-0.45
Ed<-2*p-1
SEd<-2*sqrt(p*(1-p)/N)
Ed
SEd
````


### Qn
>10\. Given the answer to 9, which of the following best describes your strategy of using a sample size of $N=25$?
>
>a. The expected value of our estimate $2\bar{X}-1$ is $d$, so our prediction will be right on.
>b. Our standard error is larger than the difference, so the chances of $2\bar{X}-1$ being positive and throwing us off were not that small. We should pick a larger sample size.
>c. The difference is 10% and the standard error is about 0.2, therefore much smaller than the difference.
>d. Because we don't know $p$, we have no way of knowing that making $N$ larger would actually improve our standard error.

Answer b. \
This is the direct consequence of the answer from the last question. The standard error is inversely proportional $\mbox{SE} \sim 1/N$ to the sample size, as stated in the answer of question 8. Therefore, the larger the sample size the smaller the standard error will be. 

## Exercises: CLT

### Qn

>1\. Write an _urn model_ function that takes the proportion of Democrats $p$ and the sample size $N$ as arguments and returns the sample average if Democrats are 1s and Republicans are 0s. Call the function `take_sample`. 

```{r}
N<-25
p<-0.45

take_sample <- function(N,p) {
  x <- sample(c(0,1), size = N, replace = TRUE, prob = c(1-p, p))
  mean(x)
}
take_sample(N,p)
```

### Qn
>2\. Now assume `p <- 0.45` and that your sample size is $N=100$. Take a sample 10,000 times and save the vector of `mean(X) - p` into an object called `errors`. Hint: use the function you wrote for exercise 1 to write this in one line of code.

```{r}
N<-100
p<-0.45
B<-10000

errors <- replicate(B, {
  X <- take_sample(N,p)
  p-mean(X)
})
qplot(errors,bins=10)
```

### Qn
>3\. The vector `errors` contains, for each simulated sample, the difference between the actual $p$ and our estimate $\bar{X}$. We refer to this difference as the _error_. Compute the average and make a histogram of the errors generated in the Monte Carlo simulation and select which of the following best describes their distributions:
>
>```{r, eval=FALSE}
>mean(errors)
>hist(errors)
>```
>
>a. The errors are all about 0.05.
>b. The errors are all about -0.05.
>c. The errors are symmetrically distributed around 0.
>d. The errors range from -1 to 1.

Answer c. \

```{r}
mean(errors)
hist(errors)
```

### Qn
>4\. The error $\bar{X}-p$ is a random variable. In practice, the error is not observed because we do not know $p$. Here we observe it because we constructed the simulation.  What is the average size of the error if we define the size by taking the absolute value $\mid \bar{X} - p \mid$ ?

```{r}
N<-100
p<-0.45
B<-10000

errors <- replicate(B, {
  X <- take_sample(N,p)
  p-mean(X)
})
mean(abs(errors))
```


### Qn
>5\. The standard error is related to the typical **size** of the error we make when predicting. We say **size** because we just saw that the errors are centered around 0, so thus the average error value is 0. For mathematical reasons related to the Central Limit Theorem, we actually use the standard deviation of `errors` rather than the average of the absolute values to quantify the typical size. What is this standard deviation of the errors?

```{r}
N<-100
p<-0.45
B<-10000
set.seed(1)

errors <- replicate(B, {
  X <- take_sample(N,p)
  p-mean(X)
})
sqrt(mean(errors^2))
```


### Qn
>6\. The theory we just learned tells us what this standard deviation is going to be because it is the standard error of $\bar{X}$. What does theory tell us is the standard error of $\bar{X}$ for a sample size of 100?

```{r}
N<-100
p<-0.45
set.seed(1)

se <- sqrt(p*(1-p)/N)
se
```


### Qn
>7\. In practice, we don't know $p$, so we construct an estimate of the theoretical prediction based by plugging in $\bar{X}$ for $p$. Compute this estimate. Set the seed at 1 with `set.seed(1)`.

```{r}
N<-100
p<-0.45
set.seed(1)

X <- take_sample(N,p)
se <- sqrt(X*(1-X)/N)
se
```

### Qn
>8\. Note how close the standard error estimates obtained from the Monte Carlo simulation (exercise 5), the theoretical prediction (exercise 6), and the estimate of the theoretical prediction (exercise 7) are. The theory is working and it gives us a practical approach to knowing the typical error we will make if we predict $p$ with $\bar{X}$. Another advantage that the theoretical result provides is that it gives an idea of how large a sample size is required to obtain the precision we need. Earlier we learned that the largest standard errors occur for $p=0.5$. Create a plot of the largest standard error for $N$ ranging from 100 to 5,000. Based on this plot, how large does the sample size have to be to have a standard error of about 1%?
>
>a. 100
>b. 500
>c. 2,500
>d. 4,000

Answer c. \
In the plot below there is a black horizontal line highlighting the 1% error border. The highest error at p=0.5 is only reached by the N values below a sample size of 2,500.

Results so far from past exercises. \
Q5 Monte Carlo simulation: 0.04949939 \
Q6 theoretical prediction: 0.04974937 \
Q7 estimate of theoretical prediction: 0.04983974 \

```{r,message=FALSE}
library(tidyverse)
N25<-replicate(100,25)
N100<-replicate(100,100)
N500<-replicate(100,500)
N1000<-replicate(100,1000)
N2500<-replicate(100,2500)
N4000<-replicate(100,4000)
N5000<-replicate(100,5000)

N<-c(N25,N100,N500,N1000,N2500,N4000,N5000)
p<-seq(0,1,length=100)
se<-sqrt(p*(1-p)/N)
df <- data.frame(p,se,N) %>% mutate(N=as.factor(N))
df %>% ggplot(aes(p,se,color=N)) + geom_point() + scale_y_continuous(breaks = seq(0, 0.1, by = 0.01)) + geom_segment(aes(x = 0, y = 0.01, xend = 1, yend = 0.01), color="black",size=0.3)

N <- seq(100, 5000, len = 100)
p <- 0.5
se <- sqrt(p*(1-p)/N)
plot(se)
````

### Qn
>9\. For sample size $N=100$, the central limit theorem tells us that the distribution of $\bar{X}$ is:
>
>a. practically equal to $p$.
>b. approximately normal with expected value $p$ and standard error $\sqrt{p(1-p)/N}$.
>c. approximately normal with expected value $\bar{X}$ and standard error $\sqrt{\bar{X}(1-\bar{X})/N}$.
>d. not a random variable.


### Qn
>10\. Based on the answer from exercise 8, the error $\bar{X} - p$ is:
>
>a. practically equal to 0.
>b. approximately normal with expected value $0$ and standard error $\sqrt{p(1-p)/N}$.
>c. approximately normal with expected value $p$ and standard error $\sqrt{p(1-p)/N}$.
>d. not a random variable.

### Qn
>11\. To corroborate your answer to exercise 9, make a qq-plot of the `errors` you generated in exercise 2 to see if they follow a normal distribution. 



### Qn
>12\. If $p=0.45$ and $N=100$ as in exercise 2, use the CLT to estimate the probability that $\bar{X}>0.5$. You can assume you know $p=0.45$ for this calculation.


### Qn
>13\. Assume you are in a practical situation and you don't know $p$. Take a sample of size $N=100$ and obtain a sample average of $\bar{X} = 0.51$. What is the CLT approximation for the probability that your error is equal to or larger than 0.01?




## Exercises: Confidence Intervals p-values


>For these exercises, we will use actual polls from the 2016 election. You can load the data from the __dslabs__ package.
>
>```{r}
>library(dslabs)
>data("polls_us_election_2016")
>```
>
>Specifically, we will use all the national polls that ended within one week before the election.
>
>```{r, message=FALSE, message=FALSE}
>library(tidyverse)
>polls <- polls_us_election_2016 %>% 
>  filter(enddate >= "2016-10-31" & state == "U.S.") 
>```

### Qn
>1\. For the first poll, you can obtain the samples size and estimated Clinton percentage with:
>
>```{r, eval=FALSE}
>N <- polls$samplesize[1]
>x_hat <- polls$rawpoll_clinton[1]/100
>```
>
>Assume there are only two candidates and construct a 95% confidence interval for the election night proportion $p$. 


### Qn
>2\. Now use `dplyr` to add a confidence interval as two columns, call them `lower` and `upper`, to the object `poll`. Then use `select` to show the `pollster`, `enddate`, `x_hat`,`lower`, `upper` variables. Hint: define temporary columns `x_hat` and `se_hat`. 

### Qn
>3\. The final tally for the popular vote was Clinton 48.2%	and Trump 46.1%. Add a column, call it `hit`, to the previous table stating if the confidence interval included the true proportion $p=0.482$ or not. 


### Qn
>4\. For the table you just created, what proportion of confidence intervals included $p$?


### Qn
>5\. If these confidence intervals are constructed correctly, and the theory holds up, what proportion should include $p$?


### Qn
>6\. A much smaller proportion of the polls than expected produce confidence intervals containing $p$. If you look closely at the table, you will see that most polls that fail to include $p$ are underestimating. The reason for this is undecided voters, individuals polled that do not yet know who they will vote for or do not want to say. 
>Because, historically, undecideds divide evenly between the two main candidates on election day, it is more informative to estimate the spread or the difference between the proportion of two candidates $d$, which in this election was $0. 482 - 0.461 = 0.021$. 
>Assume that there are only two parties and that $d = 2p - 1$, redefine `polls` as below and
>re-do exercise 1, but for the difference.
>
>```{r, message=FALSE, comment=FALSE}
>polls <- polls_us_election_2016 %>% 
>  filter(enddate >= "2016-10-31" & state == "U.S.")  %>%
>  mutate(d_hat = rawpoll_clinton / 100 - rawpoll_trump / 100)
>```



### Qn
>7\. Now repeat exercise 3, but for the difference.


### Qn
>8\. Now repeat exercise 4, but for the difference.


### Qn
>9\. Although the proportion of confidence intervals goes up substantially, it is still lower than 0.95. In the next chapter, we learn the reason for this. To motivate this, make a plot of the error, the difference between each poll's estimate and the actual $d=0.021$. Stratify by pollster.

### Qn
>10\. Redo the plot that you made for exercise 9, but only for pollsters that took five or more polls.



## Exercises: Assosiation Tests

### Qn

>1\. A famous athlete has an impressive career, winning 70% of her 500 career matches. However, this athlete gets criticized because in important events, such as the Olympics, she has a losing record of 8 wins and 9 losses. Perform a Chi-square test to determine if this losing record can be simply due to chance as opposed to not performing well under pressure.

  

### Qn
>2\. Why did we use the Chi-square test instead of Fisher's exact test in the previous exercise?
>
>a. It actually does not matter, since they give the exact same p-value.
>b. Fisher's exact and the Chi-square are different names for the same test.
>c. Because the sum of the rows and columns of the two-by-two table are not fixed so the hypergeometric distribution is not an appropriate assumption for the null hypothesis. For this reason, Fisher's exact test is rarely applicable with observational data.
>d. Because the Chi-square test runs faster.


### Qn
>3\. Compute the odds ratio of "losing under pressure" along with a confidence interval.

  
### Qn
>4\. Notice that the p-value is larger than 0.05 but the 95% confidence interval does not include 1. What explains this?
>
>a. We made a mistake in our code.
>b. These are based on t-statistics so the connection between p-value and confidence intervals does not apply.
>c. Different approximations are used for the p-value and the confidence interval calculation. If we had a larger sample size the match would be better.
>d. We should use the Fisher exact test to get confidence intervals.


### Qn
>5\. Multiply the two-by-two table by 2 and see if the p-value and confidence retrieval are a better match.


