---
title: "Ex 22 Reshaping Data"
output:
  html_document:
    df_print: paged
    number_sections: true
    toc: true
    pandoc_args: [
      "--number-sections",
      "--number-offset=22"
    ]
---



## Exercises

```{r}
library(tidyverse)
```

>1. Run the following command to define the `co2_wide` object:
>
>```{r, eval=FALSE}
>co2_wide <- data.frame(matrix(co2, ncol = 12, byrow = TRUE)) %>%
>setNames(1:12) %>%
>mutate(year = as.character(1959:1997))
>```
>
>    Use the gather function to wrangle this into a tidy dataset. Call the column with the CO2 measurements `co2` and call the month column `month`. Call the resulting object `co2_tidy`.

```{r}
str(co2)
co2_wide <- data.frame(matrix(co2, ncol = 12, byrow = TRUE)) %>% setNames(1:12) %>% mutate(year=as.character(1959:1997))
co2_tidy <- data.frame(matrix(co2, ncol = 12, byrow = TRUE)) %>% setNames(1:12) %>% mutate(year=as.factor(1959:1997)) %>% gather(key=month, value=co2, -year, convert=TRUE)
str(co2_tidy)
co2_tidy
```

>2. Plot CO2 versus month with a different curve for each year using this code:
>
>```{r, eval=FALSE}
>co2_tidy %>% ggplot(aes(month, co2, color = year)) + geom_line()
>```
>
>    If the expected plot is not made, it is probably because `co2_tidy$month` is not numeric:
>
>```{r, eval=FALSE}
>class(co2_tidy$month)
>```
>
>
>    Rewrite the call to gather using an argument that assures the month column will be numeric. Then make the plot.


```{r}
co2_tidy %>% ggplot(aes(month, co2, color = year)) + geom_line() + scale_x_continuous(breaks = c(1:12))
```

>3. What do we learn from this plot?
>
>    A. CO2 measures increase monotonically from 1959 to 1997.
>
>    B. CO2 measures are higher in the summer and the yearly average increased from 1959 to 1997.
>
>    C. CO2 measures appear constant and random variability explains the differences.
>
>    D. CO2 measures do not have a seasonal trend.

Answer A and B.

In order to see the monotonical increase of the co2 value (anwer A) and the seasonal trend over the course of each year (answer B), I've changed the class of the column year from numeric to factor, so that ggplot knows I want the lines to be connected for each year for the months.

```{r}
co2_tidy %>% filter(year==1959 | year==1969) %>% ggplot(aes(month, co2, color = year)) + geom_point() + geom_line() + scale_x_continuous(breaks = c(1:12))
```

>4. Now load the `admissions` data set, which contains admission information for men and women across six majors and keep only the admitted percentage column:
>
>```{r, eval=FALSE}
>load(admissions)
>dat <- admissions %>% select(-applicants)
>```
>
>    If we think of an observation as a major, and that each observation has two variables (men admitted percentage and women admitted percentage) then this is not tidy. Use the `spread` function to wrangle into tidy shape: one row for each major.

```{r}
library(dslabs)
data(admissions)
dat <- admissions %>% select(-applicants)
dat %>% spread(gender,admitted)
```

>5. Now we will try a more advanced wrangling challenge. We want to wrangle the admissions data so that for each major we have 4 observations: `admitted_men`, `admitted_women`, `applicants_men` and `applicants_women`.  The _trick_ we perform here is actually quite common: first gather to generate an intermediate data frame and then spread to obtain the tidy data we want. We will go step by step in this and the next two exercises.
>
>    Use the gather function to create a `tmp` data.frame with a column containing the type of observation `admitted` or `applicants`. Call the new columns `key` and value.

```{r}
tmp_gather <- admissions %>% gather(key, value, -major,-gender)
tmp_gather
```

>6. Now you have an object `tmp` with columns `major`,  `gender`, `key` and  `value`. Note that if you combine the key and gender, we get the column names we want: `admitted_men`, `admitted_women`, `applicants_men` and `applicants_women`. Use the function `unite` to create a new column called `column_name`.

```{r}
tmp_unite <- tmp_gather %>% unite(column_name, key, gender, -major, -value)
tmp_unite
```

>7. Now use the `spread` function to generate the tidy data with four variables for each major.

```{r}
tmp_unite %>% spread(column_name, value)
```

>8. Now use the pipe to write a line of code that turns `admissions` to the table  produced in the previous exercise.

```{r}
admissions %>% gather(key, value, -major,-gender) %>% unite(column_name, key, gender, -major, -value) %>% spread(column_name,value)
```

## Additional Exercises

>1. Examine the dataset `co2`. Which of the following is true:
>
>    A. `co2` is tidy data: it has one year for each row.
>
>    B. `co2` is not tidy: we need at least one column with a character vector.
>
>    C. `co2` is not tidy:  it is a matrix not a data frame.
>
>    D. `co2` is not tidy: to be tidy we would have to wrangle it to have three columns: year, month and value; then each co2 observation has a row.

Answer D.

Definition of **tidy data** from this chapter is as follows: *each row represents one observation and the columns represent the different variables available for each of these observations*

According to this defintion the time-series object does not define a tidy data set, because it is not defined wich observation belongs to which point in time. Therefore, answer D is the most likely one. The calculations below show that the object Time-Series contains 468 data points for each month for 39 years between 1959 and 1998, which are 468 points total.

In the following I will reshape the data into a tidy form. However, after reading the exercise of this chapter I found that there is a more elegant solution by using the tidyverse functions. Anyway, here is the hard way by foot to get to the tidy dataset.

```{r}
library(tidyverse)
library(dslabs)
data("co2")
str(co2)
length(co2)
(1998-1959)*12
months <- rep(1:12,39)
years <- rep(1959:1997,12)
years <- sort(years)
tab_co2 <- as.tibble(cbind(years,months,co2)) %>% arrange(years,months)
tab_co2
tab_co2 %>% ggplot(aes(years,co2,col=months)) + geom_point()
```

>2. Examine the dataset  `ChickWeight`. Which of the following is true:
>
>    A.  `ChickWeight` is not tidy: each chick has more than one row. 
>
>    B.  `ChickWeight` is tidy: each observation, here a weight, is represented by one row. The chick from which this measurement came from is one the variables.
>
>    C.  `ChickWeight` is not a tidy: we are missing the year column.
>
>    D.  `ChickWeight` is tidy: it is stored in a data frame.

Answer B.

```{r}
data(ChickWeight)
ChickWeight
str(ChickWeight)
class(ChickWeight)
ChickWeight %>% ggplot(aes(Time,weight,col=Diet)) + geom_point()
```

>3. Examine the dataset `BOD`. Which of the following is true:
>
>    A. `BOD` is not tidy: it only has six rows.
>
>    B. `BOD` is not tidy: the first column is just an index.
>
>    C. `BOD` is tidy: each row is an observation with two values, time and demand.
>
>    D. `BOD` is tidy: all small datasets are tidy by definition.

Anser C.

```{r}
data(BOD)
BOD
BOD %>% ggplot(aes(Time,demand)) + geom_line() + geom_point()
```

>4. Which of the following datasets is tidy (you can pick more than one):
>
>    A. `BJsales`
>
>    B. `EuStockMarkets`
>
>    C. `DNase`
>
>    D. `Formaldehyde`
>
>    E. `Orange`
>
>    F. `UCBAdmissions`

Answers C,D and E.

The datasets, which were plotted below are tidy datasets. These are DNase, Formaldehyde and Orange, because each observation is in a row and each column is a variable that describes the observation. The first two datasets BJsales and EuStockMarkets are time-series and need further transformation to be tidy. The dataset UCBAdmissions contains tabulated data within a table and is not tidy.
```{r}
data(BJsales)
str(BJsales)
data(EuStockMarkets)
str(EuStockMarkets)
data(DNase)
str(DNase)
DNase
DNase %>% ggplot(aes(density,conc,col=Run)) + geom_point() + geom_line()
data(Formaldehyde)
str(Formaldehyde)
Formaldehyde
Formaldehyde %>% ggplot(aes(carb,optden)) + geom_line() + geom_point()
data(Orange)
str(Orange)
Orange
Orange %>% ggplot(aes(age,circumference,col=Tree)) + geom_point() + geom_line()
data(UCBAdmissions)
str(UCBAdmissions)
UCBAdmissions
```

