---
title: "Summary 14 Probability"
output:
  html_document:
    df_print: paged
    number_sections: true
    toc: true
    pandoc_args: [
      "--number-sections",
      "--number-offset=14"
    ]
---



## Descrete Probability

>A more tangible way to think about the probability of an event is as the **proportion** of times the event occurs when we repeat the experiment an infinite number of times, independently, and under the same conditions.

Generate an urn with beads and pick a random one.
```{r}
beads <- rep(c("red", "blue"), times = c(2,3))
beads
sample(beads, 1)
```

>This line of code produces one random outcome. We want to repeat this experiment an infinite number of times, but it is impossible to repeat forever. Instead, we *repeat the experiment a large enough number of times to make the results practically equivalent to repeating forever*. This is an example of a **Monte Carlo simulation**.

```{r}
B <- 10000
events <- replicate(B, sample(beads, 1))
tab <- table(events)
tab
prop.table(tab)
```

If B gets larger the proportions get closer to 3/5=.6 and 2/5=.4.
Setting seeds to get exactly reproducable results as listed in the book.

```{r}
set.seed(1986)
events <- sample(beads, B, replace = TRUE)
prop.table(table(events))
```

